---
title: "Data Cleaning"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


Importing the Data, filter only English-language films 
```{r}
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

EDA - genre count
```{r}
genre_df = movie_df |>
  mutate(genres = str_remove_all(genres, "\\[|\\]|'")) |>
  separate_rows(genres, sep = ",\\s*")

genre_count = 
  genre_df |>
  group_by(genres) |>
  summarise(film_count = n()) |>
  arrange (desc(film_count))

print (genre_count)

```

EDA - genre df pivot wider 

```{r}
genre_wide = genre_df |> 
  mutate(
    studios = as.character(studios),
    description = as.character(description),
    genres = as.character(genres)
  ) |> 
  mutate(value = 1) |> 
  group_by(
    film_title, director, average_rating, runtime, original_language,
    description, studios, watches, list_appearances, likes, fans,
    lowest, medium, highest, total_ratings, genres
  ) |> 
  summarise(value = max(value), .groups = "drop") |> 
  pivot_wider(
    id_cols = c(
      film_title, director, average_rating, runtime, original_language,
      description, studios, watches, list_appearances, likes, fans,
      lowest, medium, highest, total_ratings
    ),
    names_from = genres,
    values_from = value,
    values_fill = 0
  ) |> 
  janitor::clean_names()
```

EDA - top 5 films based on watches (mainstream popularity)
```{r}
top_films = 
  movie_df |>
  arrange(desc(watches)) |>
  select(film_title, director, watches) |>
  slice_head(n = 5)

print(top_films)

```

EDA - top 5 films based on average rating (critical acclaim)
```{r}
top_rated = 
  movie_df |>
  arrange(desc(average_rating)) |>
  select(film_title, director, average_rating) |>
  slice_head(n = 5)

print(top_rated)

```

EDA - distribution of movie runtimes (please edit the aesthetics of it)
```{r}
ggplot(movie_df, aes(x = runtime)) +
  geom_histogram () +
  labs (title = "Distribution of Movie Runtimes",
        x = "Runtime (minutes)", y = "Count")
```

Is this an EDA? - relationship between list appearances and likes (please edit the aesthetics of it)
```{r}
ggplot(movie_df, aes(x=list_appearances, y = likes)) +
  geom_point (alpha = 0.5) +
  geom_smooth () + 
  labs (title = "Relationship between List Appearances and Likes",
        x = "List Appearances", y = "Likes")
```

Is this an EDA? - relationship between runtime and average rating (please edit the aesthetics of it)
```{r}
ggplot(movie_df, aes(x=runtime, y=average_rating)) +
  geom_point(alpha=0.5) +
  geom_smooth() +
  labs (title = "Relationship between Runtime and Average Rating",
        x = "Runtime (minutes)", y = "Average Rating")
```

Cont. EDA of relationship between runtime and average rating
```{r}
runtime_rate_df =
  movie_df |>
  mutate(runtime_category = case_when(
    runtime <= 90 ~ "Short",
    runtime > 90 & runtime <= 120 ~ "Medium",
    runtime > 120 ~ "Long"
  ))

runtime_rate_df |>
  group_by (runtime_category) |>
  summarise (avg_rating = mean(average_rating, na.rm = TRUE)) |>
  arrange (desc(avg_rating))

```

The next two chunks of code is the same, only nature of run time variable is different 


Linear Regression Model to further analyze the relationship between run time and average rating (run time as continuous variable)
```{r}
runtime_rate_model_cont = lm(average_rating ~ runtime, data = movie_df)
```

Linear Regression Model to further analyze the relationship between run time and average rating (run time as categorical variable )
```{r}
runtime_rate_model_cat = lm(average_rating ~runtime_category, data = runtime_rate_df)
```

The next two chunks of code is the same, only nature of run time variable is different 

Linear Regression Model to analyze the relationship between watches and run time (run time as cont. variable)
```{r}
runtime_watch_model_cont = lm(watches ~ runtime, data = movie_df)
```

Linear Regression Model to analyze the relationship between watches and run time (run time as categorical variable)
```{r}
runtime_watch_model_cat = lm(watches ~ runtime_category, data = runtime_rate_df)
```

