---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r include = FALSE}
# Load libraries
library(tidyverse)
library(plotly)
library(patchwork)

# Preset settings

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# Import and clean movie dataset
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

## Univariate Exploratory Analysis

To better understand our dataset, we are first going to explore the distributions for each variable of interest: average rating, runtime, genre, watches, fans, likes, and list appearances.

### Average Rating

```{r}
movie_df |> 
  plot_ly(x = ~average_rating,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Average Rating",
    xaxis = list(title = "Average Rating"),
    yaxis = list(title = "Count")
  )
```

We are interested in looking at a film's average rating as an outcome variable. The histogram representing movie average ratings follows a normal distribution, centered at around 3.25. 

### Runtime

```{r}
movie_df |> 
  plot_ly(x = ~runtime,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Runtimes",
    xaxis = list(title = "Runtime (minutes)"),
    yaxis = list(title = "Count")
  )
```

We will explore predictors possibly associated with average ratings, starting with film's runtime in minutes. The histogram follows a right skewed distribution, with the tail extending beyond 600 minutes. While the distribution is predominantly centered at around 100 minutes, there is also a spike of films that have runtimes under 20 minutes. 

### Genre

```{r}
#clean genre variable
genre_df = 
  movie_df |>
  mutate(genres = str_remove_all(genres, "\\[|\\]|'")) |>
  separate_rows(genres, sep = ",\\s*")

#summarize genre variable and only show top 18 genres
genre_clean_df = 
  genre_df |> 
  group_by(genres) |> 
  summarize(
    n = n()
  ) |> 
  arrange(desc(n)) |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) 

genre_clean_df |> 
  mutate(
    genres = fct_reorder(genres, -n)
  ) |> 
  plot_ly(x = ~genres, y = ~n, color = ~genres,
          type = "bar", colors = "viridis",
          showlegend = FALSE
  ) |> 
  layout(
    title = "Distribution of Genres",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Count")
  )
```

A bar chart was created to show the top 18 film genres represented in the dataset. We can see that drama, comedy, and thriller are the most common genres, with 3346 films classified as dramas, 2620 classified as comedy, and 1940 films classified as thriller. Since movies can be labeled as more than one genre, a  movie can contribute to multiple categories. 

### Popularity Metrics

```{r}
#create function to create plots for each variable
films_plot = function(df, xvar) {
  
    plot = 
      df |> 
      plot_ly(x = df[[xvar]],
              type = "histogram" , colors = "viridis",
              name = xvar
      ) 

  plot
  
}

#Plot for number of watches
watches_plot = 
  films_plot(movie_df, "watches") |> 
  layout(
    xaxis = list(range = c(0, 1000000))
  )

#Plot for nans
fans_plot = 
  films_plot(movie_df, "fans") |> 
  layout(
    xaxis = list(range = c(0, 50000))
  )

#Plot for number of likes
likes_plot = 
  films_plot(movie_df, "likes") |> 
  layout(
    xaxis = list(range = c(0, 500000))
  )


#Plot for number of List Appearances
list_plot = 
  films_plot(movie_df, "list_appearances") |> 
  layout(
    xaxis = list(range = c(0, 200000))
  )

#Combine plots
subplot(watches_plot, fans_plot, likes_plot, list_plot, nrows = 4) |> 
  layout(
    yaxis = list(title = "Count"),
    yaxis2 = list(title = "Count"),
    yaxis3 = list(title = "Count"),
    yaxis4 = list(title = "Count")
  )
```

We are also interested in exploring a film's popularity metrics, which includes the number of watches, fans, likes, and list appearances. The histograms show that number of watches, fans, likes, and list appearances all follow a heavily right skewed distribution, with most movies centered near the lower end of the scale. Due to the skewdness of these variables, it may be relevant to transform or categorize these variables before conducting further analyses. 

*Note*: Due to the extreme skewdness of the distributions, the plots are zoomed in to better visualize the distribution where most values lie. Users can zoom out to better understand the nature of the skewedness. 

## Correlation Matrix

```{r}
vars = 
  movie_df |> 
  select(runtime, watches, likes, fans, list_appearances)

cor_matrix =
  cor(vars, use = "complete.obs") |> 
  round(2)

plot_ly(
  z = cor_matrix, x = colnames(cor_matrix), y = colnames(cor_matrix),
  type = "heatmap", colors = "Reds"
  ) |> 
  layout(
    title = "Heat Map of Correlation Matrix"
  )
```

A heatmap representing the correlation matrix was created to better illustrate the correlation between our possible predictors and assess multicollinearity. Darker red represents a strong relationship between the pair of variables, while the lighter color represent a weaker relationship. Since the scale in the heat map demonstrates that correlatation coefficients are above, this means that all pairs of variables resemble a positive relationship. There appears to be a high collinearity between likes and watches. 

## Bivariate Exploratory Analysis

### Runtime vs Average Rating

```{r, message = FALSE}
#continuous runtime
cont_runtime_rating_plot = 
  movie_df |> 
  ggplot(aes(x = log(runtime), y = average_rating)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  labs (title = "Continuous Runtime\n vs Average Rating",
          x = "Runtime (minutes)", y = "Average Rating")

#categorical runtime
runtime_cat_df = 
  movie_df |>
  mutate(runtime_category = case_when(
    runtime < 100 ~ "Short",
    runtime >= 100 ~ "Long"
  ))

cat_runtime_rating_plot = 
runtime_cat_df |> 
  ggplot(aes(x = runtime_category, y = average_rating)) +
  geom_boxplot() +
  labs (title = "Categorical Runtime\n vs Average Rating",
          x = "Runtime Group", y = "Average Rating")

#patchwork
cont_runtime_rating_plot + cat_runtime_rating_plot
```

```{r}
#table with sample size
runtime_cat_df |>
  group_by(runtime_category) |>
  summarize(
    n = n()
  ) |>
  arrange(desc(n)) |> 
  knitr::kable()
```

### Popularity Metrics vs Average Ratings

```{r, message = FALSE, warning = FALSE}
movie_df |> 
  select(film_title, watches, likes, fans, list_appearances, average_rating) |> 
  pivot_longer(
    cols = watches:list_appearances,
    names_to = "metric",
    values_to = "value"
  ) |> 
  select(film_title, metric, value, average_rating) |> 
  mutate(
    metric = case_match(
      metric,
      "watches" ~ "Watches",
      "likes" ~ "Likes",
      "fans" ~ "Fans",
      "list_appearances" ~ "List Appearances"
    )
  ) |> 
  ggplot(aes(x = log(value), y = average_rating)) +
  geom_point(aes(color = metric), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_wrap(vars(metric), nrow = 2) +
  labs(
    x = "log(Value of Metric)",
    y = "Average Rating",
    title = "Popularity Metrics vs Average Rating"
  )
```

Popularity metrics includes number of watches, fans, likes, and list appearances.

### Average Rating per Genre

```{r}
genre_df |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) |> 
  mutate(
    genres = fct_reorder(genres, average_rating, .fun = median)
  ) |> 
  plot_ly(
    x = ~genres, y = ~average_rating, color = ~genres,
    type = "box", colors = "viridis",
    showlegend = FALSE
  )
```

## Outliers

Distributions for many variables in our dataset are skewed to the right. To explore the outliers in the dataset, we will create tables of top films based on average rating, list_appearances, and fans. 

### Top Rated Films 

```{r}
#create function for top film
top_film = function(df, variable) {
  
  table = 
    df |>
    arrange(desc({{variable}})) |>
    select(film_title, director, {{variable}}) |>
    slice_head(n = 5) |> 
    knitr::kable()

  table
  
}

top_film(movie_df, average_rating)
```

### Top Watched Films

```{r}
top_film(movie_df, watches)
```

### Top Films by List appearances

```{r}
top_film(movie_df, list_appearances)
```


