---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r include = FALSE}
# Load libraries
library(tidyverse)
library(plotly)
library(patchwork)

# Preset settings

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# Import and clean movie dataset
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

## Univariate Exploratory Analysis

### Average Rating

We are interested in looking at a films' average rating as an outcome variable. We will take a look at its distribution using a histogram. 

```{r}
movie_df |> 
  plot_ly(x = ~average_rating,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Movies' Average Rating",
    xaxis = list(title = "Average Rating"),
    yaxis = list(title = "Count")
  )
```

The histogram representing movie average ratings follows a normal distribution, centered at around 3.25. 

### Runtime

We will explore predictors possibly associated with average ratings, starting with film run time. The distribution of films' runtime is shown using a histogram below. 

```{r}
movie_df |> 
  plot_ly(x = ~runtime,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Movie Runtimes",
    xaxis = list(title = "Runtime (minutes)"),
    yaxis = list(title = "Count")
  )
```

The histogram representing movie runtimes follows a relatively normal distribution, with a slight right skew, centered at around 125 minutes. 

### Genre

Next, we will explore the distribution of film genres. We created bar chart representing the top 18 film genres that movies are associated with. 

```{r}
#clean genre variable
genre_df = 
  movie_df |>
  mutate(genres = str_remove_all(genres, "\\[|\\]|'")) |>
  separate_rows(genres, sep = ",\\s*")

#summarize genre variable and only show top 18 genres

genre_clean_df = 
  genre_df |> 
  group_by(genres) |> 
  summarize(
    n = n()
  ) |> 
  arrange(desc(n)) |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) 

genre_clean_df |> 
  mutate(
    genres = fct_reorder(genres, -n)
  ) |> 
  plot_ly(x = ~genres, y = ~n, color = ~genres,
          type = "bar", colors = "viridis",
          showlegend = FALSE
  ) |> 
  layout(
    title = "Distribution of Movie Genres",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Count")
  )
```

Based on the bar chart, drama, comedy, and thrillers are the top 3 film genres. Music, war, and western movies are less common. 

### Watches, Fans, Likes, and List Appearance

We are also interested in exploring a film's number of watches, fans, likes, and list appearances and use them as potential predictors of average rating. 

```{r}
#create function to create plots for each variable
films_plot = function(df, xvar) {
  
    plot = 
      df |> 
      plot_ly(x = df[[xvar]],
              type = "histogram" , colors = "viridis",
              name = xvar
      ) 

  plot
  
}

#Plot for number of watches
watches_plot = 
  films_plot(movie_df, "watches")

#Plot for nans
fans_plot = 
  films_plot(movie_df, "fans")

#Plot for number of likes
likes_plot = 
  films_plot(movie_df, "likes")


#Plot for number of List Appearances
list_plot = 
  films_plot(movie_df, "list_appearances")

#Combine plots
subplot(watches_plot, fans_plot, likes_plot, list_plot, nrows = 4) |> 
  layout(
    yaxis = list(title = "Count"),
    yaxis2 = list(title = "Count"),
    yaxis3 = list(title = "Count"),
    yaxis4 = list(title = "Count")
  )
```

The histograms show that number of watches, fans, likes, and list appearances all a follow heavily right skewed distribution, with most movies centered near the lower end of the scale. Due to the skewdness of these variables, it may be relevant to transform or categorize these variables before conducting further analyses. 

## Bivariate Exploratory Analysis

### Continuous runtime vs average_rating
```{r, message = FALSE}
movie_df |> 
ggplot(aes(x = log(runtime), y = average_rating)) +
  geom_point(alpha = 0.5) +
  geom_smooth(se = FALSE) +
  labs (title = "Relationship between Runtime and Average Rating",
        x = "Runtime (minutes)", y = "Average Rating")
```

### Categorical runtime vs average_rating

Boxplot

```{r}
runtime_rate_df =
  movie_df |>
  mutate(runtime_category = case_when(
    runtime <= 90 ~ "Short",
    runtime > 90 & runtime <= 120 ~ "Medium",
    runtime > 120 ~ "Long"
  ))

runtime_rate_df |>
  group_by(runtime_category) |>
  summarise(avg_rating = mean(average_rating, na.rm = TRUE)) |>
  arrange(desc(avg_rating))

runtime_rate_df |> 
  ggplot(aes(x = runtime_category, y = average_rating)) +
  geom_boxplot()


runtime_rate_df |>
  group_by(runtime_category) |>
  summarize(
    n = n()
  )

 

runtime_rate_alt2_df = 
  movie_df |>
  mutate(runtime_category = case_when(
    runtime <= 100 ~ "Short",
    runtime > 100 ~ "Long"
  ))

runtime_rate_alt2_df |> 
  ggplot(aes(x = runtime_category, y = average_rating)) +
  geom_boxplot()
  
runtime_rate_alt2_df |>
  group_by(runtime_category) |>
  summarize(
    n = n()
  ) |>
  arrange(desc(n))

```

### Watches, Fans, Likes, and List appearances vs Average Ratings

```{r, message = FALSE, warning = FALSE}
movie_df |> 
  select(film_title, watches, likes, fans, list_appearances, average_rating) |> 
  pivot_longer(
    cols = watches:list_appearances,
    names_to = "metric",
    values_to = "value"
  ) |> 
  select(film_title, metric, value, average_rating) |> 
  mutate(
    metric = case_match(
      metric,
      "watches" ~ "Watches",
      "likes" ~ "Likes",
      "fans" ~ "Fans",
      "list_appearances" ~ "List Appearances"
    )
  ) |> 
  ggplot(aes(x = log(value), y = average_rating)) +
  geom_point(aes(color = metric), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_wrap(vars(metric), nrow = 2) +
  labs(
    x = "log(Value of Metric)",
    y = "Average Rating",
    title = "Popularity Metrics vs Average Rating"
  )
```

### Average Rating per Genre

```{r}
genre_df |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) |> 
  mutate(
    genres = fct_reorder(genres, average_rating, .fun = median)
  ) |> 
  plot_ly(
    x = ~genres, y = ~average_rating, color = ~genres,
    type = "box", colors = "viridis",
    showlegend = FALSE
  )
```

## Top Films

### Top watched films 

```{r}
#create function for top film

top_film = function(df, variable) {
  
  table = 
    df |>
    arrange(desc({{variable}})) |>
    select(film_title, director, {{variable}}) |>
    slice_head(n = 5) |> 
    knitr::kable()

  table
  
}

top_film(movie_df, watches)
```

### Top rated films 

```{r}
top_film(movie_df, average_rating)
```

### Films with most fans

```{r}
top_film(movie_df, fans)
```

### Films with most likes

```{r}
top_film(movie_df, likes)
```

### Films with most list appearances

```{r}
top_film(movie_df, list_appearances)
```

Bar plots instead???

```{r}
movie_df |> 
  select(film_title, director, watches, likes, fans, list_appearances, average_rating) |> 
  pivot_longer(
    cols = watches:list_appearances,
    names_to = "metric",
    values_to = "value"
  ) |> 
  select(film_title, director, metric, value, average_rating) |> view()
```



