---
title: "Exploratory Data Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r include = FALSE}
# Load libraries
library(tidyverse)
library(plotly)
library(patchwork)

# Preset settings

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# Import and clean movie dataset
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

## Univariate Exploratory Analysis

To better understand our dataset, we are first going to explore the distributions for each variable of interest: average rating, runtime, genre, watches, fans, likes, and list appearances.

### Average Rating

```{r}
movie_df |> 
  plot_ly(x = ~average_rating,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Average Rating",
    xaxis = list(title = "Average Rating"),
    yaxis = list(title = "Count")
  )
```

We are interested in looking at a film's average rating as an outcome variable. The histogram representing movie average ratings follows a normal distribution, centered at around 3.25. 

### Runtime

```{r}
movie_df |> 
  plot_ly(x = ~runtime,
          type = "histogram", colors = "viridis"
  ) |> 
  layout(
    title = "Distribution of Runtimes",
    xaxis = list(title = "Runtime (minutes)"),
    yaxis = list(title = "Count")
  )
```

We will explore predictors possibly associated with average ratings, starting with film's runtime in minutes. The histogram follows a right skewed distribution, with the tail extending beyond 600 minutes. While the distribution is predominantly centered at around 100 minutes, there is also a spike of films that have runtimes under 10 minutes, demonstrating a presence of shorter films in the dataset.

### Genre

```{r}
#clean genre variable
genre_df = 
  movie_df |>
  mutate(genres = str_remove_all(genres, "\\[|\\]|'")) |>
  separate_rows(genres, sep = ",\\s*")

#summarize genre variable and only show top 18 genres
genre_clean_df = 
  genre_df |> 
  group_by(genres) |> 
  summarize(
    n = n()
  ) |> 
  arrange(desc(n)) |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) 

genre_clean_df |> 
  mutate(
    genres = fct_reorder(genres, -n)
  ) |> 
  plot_ly(x = ~genres, y = ~n, color = ~genres,
          type = "bar", colors = "viridis",
          showlegend = FALSE
  ) |> 
  layout(
    title = "Distribution of Genres",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Count")
  )
```

A bar chart was created to show the top 18 film genres represented in the dataset. Since movies can be labeled as more than one genre, a movie can contribute to multiple categories. Therefore, the bar charts illustrate counts of genre classification in the dataset. We can see that drama, comedy, and thriller are the most common genres, with 3346 films classified as dramas, 2620 classified as comedy, and 1940 films classified as thriller. The complexity in interpretation and the wide variety of different genres present in the dataset influenced our later modeling decision.  



### Popularity Metrics

```{r}
#create function to create plots for each variable
films_plot = function(df, xvar) {
  
    plot = 
      df |> 
      plot_ly(x = df[[xvar]],
              type = "histogram" , colors = "viridis",
              name = xvar
      ) 

  plot
  
}

#Plot for number of watches
watches_plot = 
  films_plot(movie_df, "watches") |> 
  layout(
    xaxis = list(range = c(0, 1000000), title = ""),
    yaxis = list(title = "")
  )

#Plot for fans
fans_plot = 
  films_plot(movie_df, "fans") |> 
  layout(
    xaxis = list(range = c(0, 50000), title = ""),
    yaxis = list(title = "")
  )

#Plot for number of likes
likes_plot = 
  films_plot(movie_df, "likes") |> 
  layout(
    xaxis = list(range = c(0, 500000), title = ""),
    yaxis = list(title = "")
  )

#Plot for number of List Appearances
list_plot = 
  films_plot(movie_df, "list_appearances") |> 
  layout(
    xaxis = list(range = c(0, 200000), title = ""),
    yaxis = list(title = "")
  )

#Combine plots
subplot(watches_plot, fans_plot, likes_plot, list_plot, nrows = 4) |> 
  layout(
    annotations = 
      list(
        list(
        x = 0.5, y = -0.15,
        text = "Popularity Metric", xref = "paper", yref = "paper",
        showarrow = FALSE, font = list(size = 14)
      ),
      list(
        x = -0.12, y = 0.5,
        text = "Count", textangle = -90,
        xref = "paper", yref = "paper",
        showarrow = FALSE, font = list(size = 14)
      )
      ),
    margin = list(b = 60, l = 60)
      
  )
```

We are also interested in exploring a film's popularity metrics, which includes the number of watches, fans, likes, and list appearances. The histograms show that number of watches, fans, likes, and list appearances all follow a heavily right skewed distribution, with most movies centered near the lower end of the scale. Due to the skewdness of these variables, it may be relevant to transform or categorize these variables before conducting further analyses. 

*Note*: Due to the extreme skewdness of the distributions, the plots are zoomed in to better visualize the distribution where most values lie. Users can zoom out by clicking on "Autoscale" icon on the top right of the figure to better understand the true nature of the skewedness. 

## Correlation Matrix

```{r}
#Select desired predictors 
vars = 
  movie_df |> 
  select(runtime, watches, likes, fans, list_appearances)

#Calculate correlation coefficients for each pairwise relationship
cor_matrix =
  cor(vars, use = "complete.obs") |> 
  round(2)

#Create heat map
plot_ly(
  z = cor_matrix, x = colnames(cor_matrix), y = colnames(cor_matrix),
  type = "heatmap", colors = "Reds"
  ) |> 
  layout(
    title = "Heat Map of Correlation Matrix"
  )
```

A heat map representing the correlation matrix was created to examine multicollinearity among our potential continuous predictors. Darker red represents a strong relationship between the pair of variables, while the lighter color represent a weaker relationship. Since the scale in the heat map demonstrates that correlatation coefficients are above 0, meaning that all pairs of variables resemble a positive relationship. There appears to be a high collinearity between likes and watches, with a correlation coefficnet of 0.95. To prevent multicollinearity and overfitting our model, we decided to remove likes as a potential predictor when conducting further regression analysis and model building. 


## Bivariate Exploratory Analysis

A bivariate exploratory analysis is conducted to examine relationships between our potential predictors and average rating. 

### Runtime vs Average Rating

```{r, message = FALSE}
#Continuous runtime
cont_runtime_rating_plot = 
  movie_df |> 
  ggplot(aes(x = runtime, y = average_rating)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  labs (title = "Continuous Runtime\n vs Average Rating",
          x = "Runtime (minutes)", y = "Average Rating")

#Categorical runtime
runtime_cat_df = 
  movie_df |>
  mutate(runtime_category = case_when(
    runtime < 100 ~ "Short: <100 minutes",
    runtime >= 100 ~ "Long: ≥100 minutes"
  ))

cat_runtime_rating_plot = 
runtime_cat_df |> 
  ggplot(aes(x = runtime_category, y = average_rating)) +
  geom_boxplot() +
  labs(title = "Categorical Runtime\n vs Average Rating",
          x = "Runtime", y = "Average Rating") +
  scale_x_discrete(
    labels = c(
      "Short: <100 minutes" = "Short:\n <100 minutes",
      "Long: ≥100 minutes" = "Long:\n ≥100 minutes"
    )
  )

#patchwork
cont_runtime_rating_plot + cat_runtime_rating_plot
```

```{r}
#table with sample size
runtime_cat_df |>
  group_by(runtime_category) |>
  summarize(
    n = n()
  ) |>
  arrange(desc(n)) |> 
  knitr::kable()
```

By operationalizing runtime as a continuous variable, we noticed that the relationship between runtime and average rating does not resemble a linear trend. There is a large clustering of points centered at runtime of around 100 minutes, covering a wide range of average rating, from 1 to above 4.5. There is also clustering near the lower end of runtime, mainly covering average ratings between 3 and 4.5. 

We reoperationalized runtime as a categorical variable, where film runtimes under 100 minutes are considered "short", while runtimes greater than or equal to 100 minutes are classified as "long." We decided to have the cut off at 100 minutes to maintain equal sample size per runtime group, shown in the table above. With long runtime group having a median average rating of 3.4 and the short runtime group having a median aveage rating of 3.25, there seems to be minimal difference in average rating between the two runtime groups. 

### Popularity Metrics vs Average Ratings

```{r, message = FALSE, warning = FALSE}
movie_df |> 
  select(film_title, watches, likes, fans, list_appearances, average_rating) |> 
  pivot_longer(
    cols = watches:list_appearances,
    names_to = "metric",
    values_to = "value"
  ) |> 
  select(film_title, metric, value, average_rating) |> 
  mutate(
    metric = case_match(
      metric,
      "watches" ~ "Watches",
      "likes" ~ "Likes",
      "fans" ~ "Fans",
      "list_appearances" ~ "List Appearances"
    )
  ) |> 
  ggplot(aes(x = log(value), y = average_rating)) +
  geom_point(aes(color = metric), alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(vars(metric), nrow = 2) +
  labs(
    x = "log(Value of Metric)",
    y = "Average Rating",
    title = "Popularity Metrics vs Average Rating"
  )
```

Scatterplots between different popularity metrics and average rating were also examined. Due to the skewedness of all of the popularity metric groups, those variables are plotted on the log scale. The scatterplots show that the relationship between fans, likes, list appearances, and watches on the log scale and average rating follow a positive relationship. With increase in all aspects of popularity metrics on the log scale also leads to an increase in average rating.  

### Average Rating per Genre

```{r}
genre_df |> 
  filter(genres %in% c("Action", "Adventure", "Animation", "Comedy", "Crime", "Documentary", "Drama", "Thriller", "Horror", "Romance", "Science Fiction", "Fantasy", "Mystery", "Family", "History", "Music", "War", "Western")) |> 
  mutate(
    genres = fct_reorder(genres, average_rating, .fun = median)
  ) |> 
  plot_ly(
    x = ~genres, y = ~average_rating, color = ~genres,
    type = "box", colors = "viridis",
    showlegend = FALSE
  ) |> 
  layout(
    title = "Average Rating per Genre",
    xaxis = list(title = "Genres"),
    yaxis = list(title = "Average Rating")
  )
```

A boxplot was created to visualize the average rating per genre. The boxplot shows that horror, science fiction and action have lower average ratings, with horror having the lowest median average rating of 2.96. On the other hand, a film classified as documentary, music, and history has the highest average ratings, with documentary having a median average rating of 3.59. Overall, genres with lower ratings tend to wider confidence intervals, while those with higher ratings tend to have narrower confidence intervals. Lower median average rating tend to have wider confidence intervals, while higher median average rating have narrower confidence intervals. Since confidence intervals for all genres were overlapping with one another, the differences between genres are not significant. 

## High-Performing Films

Distributions for many variables in our dataset are skewed to the right. To explore these films in the dataset, we will create tables of top films based on average rating, list_appearances, and watches 

### Top Rated Films 

```{r}
#create function for top film tables
top_film = function(df, variable) {
  
  table = 
    df |>
    arrange(desc({{variable}})) |>
    select(film_title, director, {{variable}}) |>
    slice_head(n = 5) |> 
    knitr::kable()

  table
  
}

#top rated film
top_film(movie_df, average_rating)
```

The table shown above represent the top five films based on average rating. *Radiohead: in Rainbows - From the Basement*, directed by David Barnard, has the highest average rating of 4.71. 

### Top Watched Films

```{r}
top_film(movie_df, watches)
```

The table shown above represent the top five films based on number of watches. *Barbie*, directed by Greta Gerwig, has the most watches of over 5 million watches. It is also interesting to note that two movies directed by Christopher Nolan, *Interstellar* and *The Dark Knight* are also among the most watched films.  

### Top Films by List Appearances

```{r}
top_film(movie_df, list_appearances)
```

The table shown above represent the top five films based on list appearances. *The Dark Knight*, directed by Christopher Nolan, has the most list appearances of over 800,000. Two more movies directed by Christopher Nolan, Interstellar and Oppenheimer, are also among the top films with most list appearances. It is interesting to note that both *The Dark Knight* and *Interstellar* are among the top watched films with most list appearances. 






