---
title: "Model Building"
output: 
  html_document:
    toc: true
    toc_float: true
---

This will include all code for building our model 


```{r message = FALSE}
library(tidyverse)
library(mgcv)
library(modelr)
library(glmnet)


knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

import in movie_df
```{r}
model_movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(average_rating, runtime, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```


fit a `lasso` model using `glmnet`
```{r}
x = model.matrix(average_rating ~ ., model_movie_df) [,-1]
y = model_movie_df |> pull(average_rating)
```

`lambda`
```{r}
lambda = 10^(seq(-2, 2.75, 0.1))

lasso_fit = 
  glmnet(x, y, lambda = lambda)

lasso_cv = 
  cv.glmnet(x, y, lambda = lambda)

lambda_opt = lasso_cv[["lambda.min"]]


lasso_fit_opt = glmnet(x, y, lambda = lambda_opt) 

lasso_fit_opt |> 
  broom::tidy()
  
```

Lasso model chooses the following predictors 

* runtime 
* list appearances 
* fans 
* lowest
* highest 

Using outside knowledge, we would like to include total number of watches and exclude lowest and highest (number of lowest ratings and number of highest ratings)

linear model
```{r}
linear_model = lm(average_rating ~ runtime + list_appearances + watches + fans, data = model_movie_df)


model_movie_df |> 
  add_predictions(linear_model) |> 
  add_residuals(linear_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "loess", se = FALSE)

linear_model |> broom::glance()
```


gam model
```{r}
gam_model = gam(average_rating ~ s(runtime) + s(list_appearances) + s(watches) + s(fans), data = model_movie_df)

model_movie_df |> 
  add_predictions(gam_model) |> 
  add_residuals(gam_model) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0) +
  geom_smooth(method = "loess", se = FALSE)

gam_model |> 
  broom::glance()
```

build training and testing tibbles
```{r}
cv_df = 
  crossv_mc(model_movie_df, 100) 

cv_df =
  cv_df |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))


cv_df = 
  cv_df |> 
  mutate(
    linear_mod  = map(train, \(df) lm(average_rating ~ runtime + list_appearances + watches + fans, data = df)),
    smooth_mod  = map(train, \(df) gam(average_rating ~ s(runtime) + s(list_appearances) + s(watches) + s(fans), data = df))
    ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_smooth = map2_dbl(smooth_mod, test, \(mod, df) rmse(model = mod, data = df)))


cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

we want to use a smooth model > a linear model

* smooth model has lower rmse
* also seen in the plot of resid vs. pred - the linear model is not evenly distributed across resid = 0


The optimal model that we think predicts `average_rating` is a generalized additive model containing the following predictors:

* runtime 
* list appearances 
* watches
* fans

```{r}
summary(gam_model)
```

significant p-value suggests that the non-linear relationship between the predictor and the response is statistically significant

if we want to use a linear model 
```{r}
summary(linear_model)
```

