---
title: "Regression Analysis"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r message = FALSE}
library(tidyverse)
library(mgcv)
library(modelr)
library(glmnet)
library(plotly)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```

Import data and organize genre 
```{r}
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

```{r}
genre_df = movie_df |>
  mutate(genres = str_remove_all(genres, "\\[|\\]|'")) |>
  separate_rows(genres, sep = ",\\s*")
```

```{r}
genre_wide = genre_df |> 
  mutate(
    studios = as.character(studios),
    description = as.character(description),
    genres = as.character(genres)
  ) |> 
  mutate(value = 1) |> 
  group_by(
    film_title, director, average_rating, runtime, original_language,
    description, studios, watches, list_appearances, likes, fans,
    lowest, medium, highest, total_ratings, genres
  ) |> 
  summarise(value = max(value), .groups = "drop") |>
  pivot_wider(
    id_cols = c(
      film_title, director, average_rating, runtime, original_language,
      description, studios, watches, list_appearances, likes, fans,
      lowest, medium, highest, total_ratings
    ),
    names_from = genres,
    values_from = value,
    values_fill = 0
  ) |> 
  janitor::clean_names()
```

# Linear Models

## Runtime vs. Average Rating
```{r}

```

## Runtime vs. Average Rating, adjusted for covariates
```{r}

```

## Appearances in User-Curated Lists vs. Average Rating
```{r}
movie_df |> 
  ggplot(aes(list_appearances, average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Appearances",
    y = "Average Rating",
    title = "Linear Regression of Appearances on User-Curated Lists vs. Average Rating"
  )

appearances_model = lm(average_rating ~ list_appearances, data = movie_df)

appearances_model |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(p.value = formatC(p.value, format = "e", digits = 3)) |> 
  knitr::kable()
```

The coefficient for `list_appearances` is 3.39-e06, which represents the estimated increase in rating for each unit increase in appearances in user-curated lists.

The model's p-value is 3.44e-288, which indicates that the number of times the movie appears in user-curated lists is extremely statistically significant in predicting its average rating.

## Total Watches vs. Average Rating
```{r}
movie_df |> 
  ggplot(aes(watches, average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Total Watches",
    y = "Average Rating",
    title = "Linear Regression of Total Watches vs.Average Rating"
  )

watches_model = lm(average_rating ~ watches, data = movie_df)

watches_model |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |>
  mutate(p.value = formatC(p.value, format = "e", digits = 3)) |> 
  knitr::kable()
```

The coefficient for `watches` is 3.04-e07, which represents the estimated increase in rating for each unit increase in the total number of times the movie have has watched by users.

The model's p-value is 7.41e-154, which indicates that the number of times the movie has been watched is extremely statistically significant in predicting its average rating.

## Fans vs. Average Rating
```{r}
movie_df |> 
  ggplot(aes(fans, average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  scale_x_log10() +
  labs(
    x = "Number of Fans (Log Scale)",
    y = "Average Rating",
    title = "Linear Regression of Number of Fans (Log Scale) vs. Average Rating"
  )

fans_model = lm(average_rating ~ fans, data = movie_df) 

summary(fans_model)

fans_model |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(p.value = formatC(p.value, format = "e", digits = 3)) |> 
  knitr::kable()
```

Due to `fans` being extremely skewed and having a wide value range, we used log scale to reduce extreme outliers and depict any linear relationships.

The coefficient for `fans` is 1.38-e05, which represents the estimated increase in rating of the movie for each unit increase in the number of users who have marked themselves as a fan of the film.

The model's p-value is 2.88e-126, which indicates that the number of fans is is extremely statistically significant in predicting its average rating.

### What's with the extreme numbers?

You might be wondering why both the coefficients and p-values are so small. First, consider the nature of our predictors. The number of appearances in user-curated lists, watches, and fans can easily reach the hundreds of thousands. For example, 100,000 more appearances predicts an increase of 0.339 in a film's ratings. 

Similarly, our p-values are so large due to the size of our dataset (n = 8074). Thus, even very small effect sizes can lead to statistically significant p-values. To investigate the actual effect sizes, let's look at the R-squared values of each model:

```{r}
r2_table = tibble(
 r2_model = c("Appearances in User-Curated Lists", "Watches", "Fans"),
r2 = c(summary(appearances_model)$r.squared,
summary(watches_model)$r.squared,
summary(fans_model)$r.squared
)
)

knitr::kable(r2_table, digits = 4)
```
Appearances: This model explains about 15%  of the variation in movie ratings, suggesting that more appearances in user-curated lists are predictive of a higher rating.

Watches: This model explains about 8.3% of the variation in movie ratings. While a weaker effect size, this suggests that total watches is modestly predictive of a higher rating.

Fans: This model explains about 6.8% of the variation in movie ratings. The number of fans of a film has the smallest effect on a movie rating.

Out of these variables, appearances in user-curated lists are the most predictive of higher ratings. This may be explained by users being pushed films by an algorithm that is tailored to their past watch history, liked movies, and other collected data. Thus, viewers enjoying a film catered to their preferences are likely to assign it a high rating.

## Cross-validation: Average Rating

```{r}
cv2_df = 
  crossv_mc(movie_df, 100)

cv2_df = cv2_df %>% 
  mutate(
    appearances_mod = map(.x = train, ~lm(average_rating ~ list_appearances, data = .x)),
    watches_mod = map(.x = train, ~lm(average_rating ~ watches, data = .x)),
    fans_mod = map(.x = train, ~lm(average_rating ~ fans, data = .x))
    )|> 

  mutate(
    rmse_appearances = map2_dbl(.x = appearances_mod, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_watches = map2_dbl(.x = watches_mod, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_fans = map2_dbl(.x = fans_mod, .y = test, ~rmse(model = .x, data = .y))
  )

cv2_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Root Mean Squared Errors Distributions for Average Rating"
  )

cv2_df |> 
  summarise(
    mean_appearances = mean(rmse_appearances),
    mean_watch = mean(rmse_watches),
    mean_fans = mean(rmse_fans)
  )
```

