---
title: "Final Report"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

### Motivation

We are a group of passionate movie enthusiasts who are eager to dive into the Letterboxd dataset to uncover fascinating insights about films. In an era dominated by short-form content and claims of shorter attention spans, we are curious to see if shorter movies tend to receive higher ratings. We also want to explore the secret formulas behind top-rated films: Do certain directors excel when they stick to a specific genre or keep their films under 90 minutes? Beyond that, we are excited to analyze trends in genre preferences and discover what kind of film would captivate audiences.

### Initial Questions

Through our analysis, we are trying to understand what factors contribute to higher ratings of movies. We were initially drawn to this dataset because of the variety of variables it provided. This Letterboxd dataset contained information about movie runtime, movie genre, directors, and a variety of measurements of movie rating and perceived success. We initially were very interested in answering the question of  are films directed by more "famous" directors more likely to have higher ratings? And how does genre play into this? Many directors tend to stick to their niche, and we were curious to see how their films performed when they stayed in their niche or not. We also noted that the dataset contained information about film and production studios, and we wanted to know how ratings would change dependent on bigger more well known studies and smaller lesser known ones. 

Unfortunately this idea met some challenges. Despite our dataset being robust, variables like genre, studio, and director had multiple entries for each movie and we ran into challenges separating them, and most importantly conducting regression analysis on it. With this in mind, we decided to shift our research question to look more at how certain continuous variables contribute to average rating. 

As we did this, we explored how each variable individually was associated with average rating, and shifted our question and curiousity to if these variables together can come together to create a prediction model that is able to accurately estimate the rating of a film based on factors like movie run time, number of watches, the number of times it appears on Letterboxd, and how many watchers identify as fans of the film. 

### Data

The Letterboxd dataset was downloaded from Kaggle, an online data science community that provides free public datasets for solving real-world problems. The dataset was downloaded in CSV format and uploaded to a shared GitHUb repository. The original downloaded dataset contained metadeta for 10,002 movies with 16 column variables and was 'clean' in the sense that it had no missing values. However, for the purpose of our analysis, we filtered the dataset to include only English-language films. 

During the data cleaning and preparation phase, we developed subsets of ready-to-use code from the main dataframe. This was done for two primary reasons:

1. To provide code snippets that would facilitate and expedite the subsequent exploratory data analysis and regression.

2. We anticipated potential issues with the nature of certain variables which could complicate their use as predictors in our regression model. 

Hence, this data cleaning and preparation stage involved extensive experimentation with various variables to determine which ones could be effectively coded and would run properly. Additionally, code snippets that treated variables as either categorical or continuous were provided to enable immediate application in subsequent analyses. As mentioned earlier, we discovered that some variables such as genre, director and studio were more challenging to separate and restructure. While we managed to successfully separate genre, we struggled to establish an appropriate method for handling the director and studio variables. Therefore, we opted to tweak the focus of our research questions and exclude director and studio from our analyses. 

### Exploratory Analysis

```{r include = FALSE}
# Load libraries
library(tidyverse)
library(plotly)
library(patchwork)
library(mgcv)
library(modelr)
library(glmnet)

# Preset settings

knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = .8,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

# Import and clean movie dataset
movie_df = read.csv("data/letterbox_movie_classification_dataset.csv") |>
  janitor::clean_names() |>
  filter(original_language == "English") |>
  select(film_title, director, average_rating, genres, runtime, original_language, description, studios, watches, list_appearances, likes, fans, lowest, medium, highest, total_ratings)
```

### Model Building

### Additional Analysis: Linear Models 

Across several regression analyses, we wanted to assess how individual movie characteristics of the model relate to a film’s average rating on Letterboxd. We first examined `runtime` as a predictor and found that longer movies were statistically associated with higher ratings. However, upon examining the resulting plot, there was a suggested weak linear trend. Even when applying a log-transformation or dichotomizing `runtime` into short (<100 minutes) and long (≥100 minutes) films, `runtime` remained a difficult variable to model on its own. This suggests that, individually, `runtime` provides limited insight into what drives audience reception.

```{r, warning = FALSE, message = FALSE}
appearances_plot = movie_df |> 
  ggplot(aes(list_appearances, average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Appearances",
    y = "Average Rating",
    title = "Linear Relationship Betweem Appearances on User-Curated Lists 
and Average Rating"
  )

watches_plot = movie_df |> 
  ggplot(aes(watches, average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "Total Watches",
    y = "Average Rating",
    title = "Linear Relationship Between Total Watches and Average Rating"
  )

fans_plot = movie_df |> 
  ggplot(aes(log(fans), average_rating)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    x = "log(Fans",
    y = "Average Rating",
    title = "Linear Relationship Between Number of log(fans) and Average Rating"
  )

appearances_plot / watches_plot / fans_plot
```


Next, we explored alternative predictors that reflect audience engagement, appearances on user-curated lists `list_appearances`, total `watches`, and number of `fans`. All three variables were highly statistically significant predictors of average rating and were driven by the very large sample size. Effect sizes also varied, with appearances in user-curated lists explaining the greatest proportion of variation in ratings (R² ≈ 0.15), followed by total watches (R² ≈ 0.083) and fans (R² ≈ 0.068). These findings indicate movies that are more widely circulated and recommended within the platform tend to earn higher ratings which is possibly due to how many platforms make individualized recommendations for its viewers. 

```{r, message = FALSE}
runtime_adj_model = lm(
  average_rating ~ runtime + watches + list_appearances + fans,
  data = movie_df
)

runtime_adj_model |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |>
  mutate(p.value = formatC(p.value, format = "e", digits = 3)) |> 
  knitr::kable()
```

```{r, message = FALSE}
runtime_model = lm(average_rating ~ runtime, data = movie_df)
appearances_model = lm(average_rating ~ list_appearances, data = movie_df)
watches_model = lm(average_rating ~ watches, data = movie_df)
fans_model = lm(average_rating ~ fans, data = movie_df) 
runtime_adj_model = lm(
  average_rating ~ runtime + watches + list_appearances + fans,
  data = movie_df
)

r2_table = tibble(
 r2_model = c("Runtime", "Appearances in User-Curated Lists", "Watches", "Fans", "Runtime (adjusted)"),
r2 = c(summary(runtime_model)$r.squared,
summary(appearances_model)$r.squared,
summary(watches_model)$r.squared,
summary(fans_model)$r.squared,
summary(runtime_adj_model)$r.squared
))

knitr::kable(r2_table, digits = 4)
```

Finally, incorporating multiple predictors into an adjusted model resulted in improved model performance, with the adjusted runtime model explaining approximately 19% of the variance in average ratings, the strongest result among all models evaluated. 

```{r, message = FALSE}
cv2_df = 
  crossv_mc(movie_df, 100)

cv2_df = cv2_df %>% 
  mutate(
    appearances_mod = map(.x = train, ~lm(average_rating ~ list_appearances, data = .x)),
    watches_mod = map(.x = train, ~lm(average_rating ~ watches, data = .x)),
    fans_mod = map(.x = train, ~lm(average_rating ~ fans, data = .x)),
    runtime_mod = map(.x = train, ~lm(average_rating ~ runtime, data = .x)),
    runtime_adj_mod = map(.x = train, ~lm(
  average_rating ~ runtime + watches + list_appearances + fans,
  data = .x))
    ) |> 

  mutate(
    rmse_appearances = map2_dbl(.x = appearances_mod, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_watches = map2_dbl(.x = watches_mod, .y = test, ~rmse(model = .x, data = .y)), 
    rmse_fans = map2_dbl(.x = fans_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_runtime = map2_dbl(.x = runtime_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_runtime_adj = map2_dbl(.x = runtime_adj_mod, .y = test, ~rmse(model = .x, data = .y))
  )

cv2_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Root Mean Squared Errors Distributions for Average Rating"
  )
```


Cross-validation further supported this conclusion, as the covariate-adjusted runtime model yielded the lowest RMSE. This indicated the highest predictive accuracy. Altogether, while `runtime` alone is a weak predictor of perceived movie quality, audience engagement metrics, particularly curated list appearances, emerge as more powerful indicators of how viewers rate the movie. 

### Discussion

For our additional analyses, our findings suggest that while several characteristics are statistically associated with a movie's rating, `runtime` alone showed a significant but very weak association with higher ratings and lacked a clearly linear visual pattern. In contrast, appearances on user-curated lists `list_appearances`, total `watches`, and `fans` all explained a much greater share of the variance, suggesting that movies with higher visibility and stronger community traction tend to receive higher ratings. The adjusted model that combined all the variables mentioned performed the best, which supports our idea that the more popular a film is the better ratings it will have. 

However, our analysis has limitations which include effect sizes being on the smaller side and statistical significance being driven by a large sample size. There is the possibility that these predictors may be a source of multicollinearity with underlying factors such as genre, marketing, or release year, and the dataset includes only English-language films. All these factors limit generalizability across global cinema and other genres of media. Future work could incorporate additional film attributes and nonlinear modeling techniques to better capture what contributes to audience reception and rating behavior.

=======
